---
title: "Good data programming solution"
author: "Ken Mwai"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    number_sections: false
    theme: spacelab
    highlight: tango
includes:
 css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Data loading {.tabset .tabset-fade .tabset-pills}
#### Task 1

::: {.task}
- Create a project folder with (data, code ,result , ... folders) 

- Use the mtcars dataset within R and check the structure

  - Take a first look at the data. Useful functions are  
   `dim()`, `head()`, `str()` and `summary()`.
   
- Remember to use good programming approach
:::

#### Step 1

- Create the data,results, code folders
- Create an r_script and save it in the code folder

#### Step 2 - R Script

```{r step2}
#----------------------------------------------
## f_StaR Good programming skills
## Author: Ken Mwai - April 2022
#----------------------------------------------
#----------------------------------------------
# 0 - Load libraries 
#----------------------------------------------
library(stats)
library(dplyr)
#----------------------------------------------
# 1 - Source Data
#----------------------------------------------
df1 <- mtcars

#----------------------------------------------
#' 2 - Dimension of the the data
#' Shows the number of rows and columns
#----------------------------------------------
dim(df1)

#----------------------------------------------
#' 3 - The first few observations of the data
#----------------------------------------------
head(df1)

#----------------------------------------------
#' 3 - The data structure
#----------------------------------------------
glimpse(df1)
str(df1)

#----------------------------------------------
#' 3 - The summary statistics of the data
#----------------------------------------------
summary(df1)
```
